---
title: "R Notebook"
output: html_notebook
fig_width: 10
fig_height: 10
---

```{r}
knitr::opts_chunk$set(fig.width=22, fig.height=20) 
library(tidyverse)
library(caret)
library(mgcv)
library(visreg)

```

## Import
Importing Cleaned Data
```{r}
ames = read_csv('./data/train_clean.csv')
test = read_csv('./data/test_clean.csv')
ames = ames[,order(colnames(ames))]
ames = ames %>% dplyr::select(-X1)
ames = ames[,order(colnames(ames))] %>% 
  rename('FirstFlrSF' = "1stFlrSF", 'SecFlrSF' = '2ndFlrSF', 'ThreeSeaPorch' = '3SsnPorch')


test = test[,order(colnames(test))] %>% 
  rename('FirstFlrSF' = "1stFlrSF", 'SecFlrSF' = '2ndFlrSF', 'ThreeSeaPorch' = '3SsnPorch')

test[is.na(test)] <- 0
#ames %>% mutate(PriceRange = ifelse(Neighborhood %in% medians$Neighborhood, factor(medians$PriceRange),0)) %>% select(c(Neighborhood,PriceRange))

medians = ames %>% group_by(Neighborhood) %>% summarise(MedianPrice = mean(SalePrice))
medians$PriceRange = medians$MedianPrice %>% cut_number(5, labels=c('Cheap','LowerMiddle','Middle','UpperMiddle','Expensive'))


ames = ames %>% left_join(., medians[,c('Neighborhood','PriceRange')], by='Neighborhood')
test= test %>% left_join(., medians[,c('Neighborhood','PriceRange')], by='Neighborhood')
```


## Split into train and validation
```{r}
set.seed(3)
train.idx = sample(1:nrow(ames), 8*nrow(ames)/10)
ames_train = ames[train.idx,]
ames_test = ames[-train.idx,]

```


```{r}
x_train = 
grid = 10^seq(5, -2, length = 100)
set.seed(0)
fit = cv.gamsel(x_train, y_train)#, lambda = grid)
```

```{r}
## data
set.seed(42)
x <- seq(0, 1, length=500)
y <- sin(2*pi*x)^2 + rnorm(500, sd=0.2)
df <- data.frame(x = x, y = y)

## Generate GCV traces
lambda <- 1e-6
gcv <- numeric(length = 100)

## loop over values of lambda, fixing sp
for (i in seq_along(gcv)) {
    m <- gam(y ~ s(x, k = 20), data = df, method = "GCV.Cp", sp = lambda)
    gcv[i] <- m$gcv.ubre
    lambda <- lambda * 1.1
}

## gather results
res <- data.frame(lambda = 1e-6 * 1.1^{0:99}, gcv = gcv)

## plot
ggplot(res, aes(x = lambda, y = gcv)) +
    geom_line() +
    scale_x_log10() +
    xlab(expression(lambda)) +
    ylab("Score")
```


```{r}
ames.gam <- mgcv::gam(log(SalePrice) 
                ~ s(GrLivArea, by=PriceRange, bs='cr', id=1) # Above Ground SF
                + s(TotalBsmtSF,bs='cr') # Bsmt SF
                + s(OverallQual, bs='cr', by=PriceRange) 
                + s(YearBuilt, bs='cr') 
                + Fireplaces:PriceRange
                + s(GarageArea, by = GarageFinish, bs='cr') # Garage Predictors=
                + s(MSSubClass, bs='cr')
                #+ AfterWW2 
                #+ Neighborhood:ExterCond
                #+ ti(MSSubClass, GrLivArea) 
                #+ s(HouseStyle, by= AfterWW2) # doesn't work
                + Neighborhood:OverallCond #+ BsmtQual:TotalBsmtSF + ExterCond:(sqrt(GarageArea)+sqrt(LotArea))
             #   + s(LotArea) + s(LotFrontage) +s(OpenPorchSF) # + AfterWW2 + Neighborhood, 
                ,method='ML',data=ames, select = TRUE)

ames.lm <- lm(log(SalePrice) 
                ~ sqrt(GrLivArea) #+ sqrt(TotalBsmtSF) 
              + OverallQual 
              + YearBuilt 
              + Fireplaces 
              + TotalPorchSF 
              + MSSubClass 
              + PriceRange
              
              #+ LotArea + LotFrontage +OpenPorchSF # + AfterWW2 + Neighborhood, 
                ,data=ames)

true = ames_test[,'SalePrice'][[1]]
gam.predictions = predict.gam(ames.gam, newdata = ames_test, type = 'response')
gam.errors = gam.predictions - log(true)
gam.diff =exp(gam.predictions) - true


lm.predictions = predict(ames.lm, newdata = ames_test)
lm.errors = lm.predictions - log(true)
lm.diff = exp(lm.predictions) - true


plot(density(gam.errors), col ='red')
lines(density(lm.errors), col='black', lty=2)
#plot(ames.gam)
legend("topright",legend=c('GAM Residuals','LM Residuals'),
       col=c("red","black"), lty=1:2, cex=0.8)


print(paste('GAM RMSE:',sqrt(mean(gam.errors^2))))
print(paste('LM RMSE:',sqrt(mean(lm.errors^2))))

print(paste('GAM: Error in $ as Predicted:',round(mean(abs(gam.diff)),2)))
print(paste('LM: Error in $ as Predicted:',round(mean(abs(lm.diff)),2)))

#plot(ames_test[,'SalePrice'][[1]], gam.errors)

plot(true, exp(gam.predictions), ylim=c(0,7e5), xlim=c(0,7e5))
abline(a=0,b=1)

#points(ames_test_[,'SalePrice'][[1]], exp(lm.predictions))
plot(true, exp(lm.predictions), ylim=c(0,7e5), xlim=c(0,7e5))
abline(a=0,b=1)
```

```{r}
summary(ames.gam)
```

```{r}
plot(ames.gam, se=2)
```

```{r}
visreg(ames.gam)
```


```{r}

gam.test_predictions = exp(predict.gam(ames.gam, newdata = test, type = 'response'))
submission = data.frame(gam.test_predictions)
colnames(submission) = 'SalePrice'
submission = tibble::rowid_to_column(submission,'Id')
rownames(submission) = 1461:(nrow(submission)+1460)
gm_submission = mutate(submission, Id = Id + 1460)

lm.test_predictions = exp(predict(ames.lm, newdata = test, type = 'response'))
lm_submission = data.frame(lm.test_predictions)
colnames(lm_submission) = 'SalePrice'
lm_submission = tibble::rowid_to_column(lm_submission,'Id')
rownames(lm_submission) = 1461:(nrow(lm_submission)+1460)
lm_submission = mutate(lm_submission, Id = Id + 1460)
#write.csv(gam.test_predictions,'./data/price_predictions.csv')
```

```{r}
write.table(gm_submission,file="./data/gm_predictions.csv",col.names = c("Id","SalePrice"),sep = ",",row.names = F)
write.table(lm_submission,file="./data/lm_predictions.csv",col.names = c("Id","SalePrice"),sep = ",",row.names = F)

write.table((lm_submission+gm_submission)/2,file="./data/avg_predictions.csv",col.names = c("Id","SalePrice"),sep = ",",row.names = F)
```

```{r}
which(is.na(gm_submission)) - 1460
which(is.na(lm_submission)) - 1460
length(submission)
```
```{r}

ames %>% 
  ggplot(aes(x = reorder(Neighborhood,SalePrice, median), y = SalePrice)) + 
  geom_boxplot() +
  theme_bw() +
  theme(axis.text.x=element_text(angle=45, hjust=1)) 

ames %>% 
  ggplot(aes(x = reorder(Neighborhood,SalePrice, mean), y = SalePrice)) + 
  geom_boxplot() +
  theme_bw() +
  theme(axis.text.x=element_text(angle=45, hjust=1)) +
  
  labs(x='Neighborhood', y = 'SalePrice')

```

```{r}
ames$PriceRange =ames$SalePrice %>% cut_number(3, labels=c('Cheap','Average','Expensive'))
```
```{r}
n <- 10; nn <- 100
g <- factor(round(n * runif(n * nn)))
x <- rnorm(n * nn) + sqrt(as.numeric(g))
xg <- split(x, g)
boxplot(xg, col = "lavender", notch = TRUE, varwidth = TRUE)
sapply(xg, length)
sapply(xg, mean)
```

```{r}
ames %>% 
  ggplot(aes(x=GrLivArea, y = SalePrice, col=PriceRange)) + 
  geom_point()
```


```{r}
ames %>% 
  ggplot(aes(x = reorder(Neighborhood,SalePrice, mean), y = SalePrice)) + 
  geom_boxplot() +
  theme_bw() +
  theme(axis.text.x=element_text(angle=45, hjust=1)) +
  
  labs(x='Neighborhood', y = 'SalePrice')
```

## Diagnositcs
```{r}
#plot(ames_test[,'SalePrice'][[1]], gam.errors)

plot(true, exp(gam.predictions), ylim=c(0,7e5), xlim=c(0,7e5))
abline(a=0,b=1)

#points(ames_test_[,'SalePrice'][[1]], exp(lm.predictions))
plot(true, exp(lm.predictions), ylim=c(0,7e5), xlim=c(0,7e5))
abline(a=0,b=1)

```

```{r}

plot(ames.gam,scale=0,se=1,shade=TRUE)
visreg(ames.gam)
```

```{r}
library(lattice)



visreg(ames.gam,'TotalBsmtSF', by='BsmtQual', breaks=6)
visreg(ames.gam,'GrLivArea', by='OverallQual', breaks=10)
visreg(ames.gam,'TotalBsmtSF', by='BsmtQual', overlay=TRUE, breaks=3)
#visreg(ames.gam,'Neighborhood', type='conditional')
```


```{r}
plot(ames_train_$BsmtQual, ames_train_$SalePrice)
plot(ames_train_$TotalBsmtSF, ames_train_$SalePrice)
plot(ames_train_$BsmtQual*ames_train_$TotalBsmtSF, ames_train_$SalePrice)
```

```{r}
#plot(ames_train$Neighborhood, ames_train$SalePrice)
plot(ames_train$OverallCond, ames_train$SalePrice)
plot(ames_train$OverallCond*ames_train$Neighborhood, ames_train$SalePrice)
```

```{r}
summary(ames.gam)
```


```{r}

min_n = 20
# ames_train %>%
#   group_by(Neighborhood, OverallCond) %>%
#   tally() %>%
#   filter(n > min_n) %>%
#   ggplot(aes(x=Neighborhood, y=OverallCond)) +
#   geom_point() +
#   geom_text(color='red',size=4,aes(y=OverallCond+0.2, label=n)) +
#   theme_bw() +
#   theme(axis.text.x=element_text(angle=45, hjust=1))
#
# ames_train %>%
#   group_by(MSSubClass, OverallQual) %>%
#   tally() %>%
#   filter(n > min_n) %>%
#   ggplot(aes(x=factor(MSSubClass), y=OverallQual)) +
#   geom_point() +
#   geom_text(color='red',size=4,aes(y=OverallQual+0.2, label=n)) +
#   theme_bw() +
#   theme(axis.text.x=element_text(angle=45, hjust=1))
#
#
# ames_train %>%
#   group_by(HouseStyle, OverallQual) %>%
#   tally() %>%
#   filter(n > min_n) %>%
#   ggplot(aes(x=factor(HouseStyle), y=OverallQual)) +
#   geom_point() +
#   geom_text(color='red',size=4,aes(y=OverallQual+0.2, label=n)) +
#   theme_bw() +
#   theme(axis.text.x=element_text(angle=45, hjust=1))

# Zoning relationship wrt Overall Quality of the Home
# Floating Village Residential has score >= 6
# Commercial zones have scores <= 6
# Low density residential spans all scores
# medium density tends towards higher scores

# ames_train %>%
#   group_by(MSZoning, OverallQual) %>%
#   tally() %>%
#   filter(n > min_n) %>%
#   ggplot(aes(x=factor(MSZoning), y=OverallQual)) +
#   geom_point() +
#   geom_text(color='red',size=4,aes(y=OverallQual+0.2, label=n)) +
#   theme_bw() +
#   theme(axis.text.x=element_text(angle=45, hjust=1))
#
# ames_train %>%
#   group_by(SaleType, OverallQual) %>%
#   tally() %>%
#   filter(n > min_n) %>%
#   ggplot(aes(x=factor(SaleType), y=OverallQual)) +
#   geom_point() +
#   geom_text(color='red',size=4,aes(y=OverallQual+0.2, label=n)) +
#
#   theme_bw() +
#   theme(axis.text.x=element_text(angle=45, hjust=1))
#
ames %>%
  group_by(SaleType, OverallCond) %>%
  tally() %>%
  filter(n > min_n) %>%
  ggplot(aes(x=factor(SaleType), y=OverallCond)) +
  geom_point() +
  geom_text(color='red',size=4,aes(y=OverallCond+0.2, label=n)) +
  theme_bw() +
  theme(axis.text.x=element_text(angle=45, hjust=1))

ames %>%
  group_by(GarageType, GarageCond) %>%
  tally() %>%
  filter(n > min_n) %>%
  ggplot(aes(x=factor(GarageType), y=GarageCond)) +
  geom_point() +
  geom_text(color='red',size=4,aes(y=GarageCond+0.2, label=n)) +
  theme_bw() +
  theme(axis.text.x=element_text(angle=45, hjust=1))

ames %>%
  group_by(Neighborhood, OverallQual) %>%
  tally() %>%
  filter(n > min_n) %>%
  ggplot(aes(x=reorder(Neighborhood,OverallQual,max), y=OverallQual)) +
  geom_point() +
  geom_text(color='red',size=4,aes(y=OverallQual+0.2, label=n)) +
  theme_bw() +
  theme(axis.text.x=element_text(angle=45, hjust=1))

ames %>%
  group_by(Neighborhood, OverallCond) %>%
  tally() %>%
  filter(n > min_n) %>%
  ggplot(aes(x=reorder(Neighborhood,OverallCond,max), y=OverallCond)) +
  geom_point() +
  geom_text(color='red',size=4,aes(y=OverallCond+0.2, label=n)) +
  theme_bw() +
  theme(axis.text.x=element_text(angle=45, hjust=1))

ames %>%
  group_by(OverallCond, OverallQual) %>%
  tally() %>%
  filter(n > min_n) %>%
  ggplot(aes(x=OverallCond, y=OverallQual)) +
  geom_point() +
  geom_text(color='red',size=4,aes(y=OverallQual+0.2, label=n)) +
  theme_bw() +
  theme(axis.text.x=element_text(angle=45, hjust=1))

ames %>%
  group_by(OverallQual, PriceRange) %>%
  tally() %>%
  filter(n > min_n) %>%
  ggplot(aes(y=OverallQual, x = PriceRange)) +
  geom_point() +
  geom_text(color='red',size=4,aes(y=OverallQual+0.5, label=n)) +
  theme_bw() +
  theme(axis.text.x=element_text(angle=45, hjust=1))

ames %>%
  group_by(OverallCond, PriceRange) %>%
  tally() %>%
  filter(n > min_n) %>%
  ggplot(aes(y=OverallCond, x = PriceRange)) +
  geom_point() +
  geom_text(color='red',size=4,aes(y=OverallCond+0.5, label=n)) +
  theme_bw() +
  theme(axis.text.x=element_text(angle=45, hjust=1))

#ames$PriceRange =ames$SalePrice %>% cut_number(5, labels=c('Cheap','Lower Middle', 'Middle','Upper Middle','Expensive'))



#ames %>% mutate(PriceRange = ifelse(Neighborhood %in% medians$Neighborhood, factor(medians$PriceRange),0)) %>% select(c(Neighborhood,PriceRange))
#ames = ames %>% left_join(., medians[,c('Neighborhood','PriceRange')], by='Neighborhood')

ames %>% 
  group_by(Neighborhood, PriceRange, SalePrice) %>% 
  summarise(med = median(SalePrice), n = n()) %>% 
  filter(n > min_n) %>% 
  ggplot(aes(x=reorder(Neighborhood,n,max), y = PriceRange)) + 
  geom_point(color='white') +
  geom_text(color='red',size=4,aes(y=PriceRange, label=n)) +
  theme_bw() +
  theme(axis.text.x=element_text(angle=45, hjust=1)) 



ames %>% 
  ggplot(aes(x=PriceRange, y = SalePrice)) + 
  geom_boxplot() +
  theme_bw() +
  theme(axis.text.x=element_text(angle=45, hjust=1)) 

```
```{r}

stat_box_data <- function(y, upper_limit = max(ames$SalePrice) * 1.15) {
  return( 
    data.frame(
      y = 0.98 * upper_limit,
      label = paste('count\n', length(y), '\n')
    )
  )
}

ames %>% 
  ggplot(aes(x=reorder(SaleType,SalePrice,median), y = SalePrice)) + 
  geom_boxplot() +
  stat_summary(
    fun.data = stat_box_data, 
    geom = "text", 
    hjust = 0.5,
    vjust = 0.9
  ) + 
  theme_bw() +
  labs(x = 'Sale Type', y = 'Sale Price')
  theme(axis.text.x=element_text(angle=45, hjust=1)) 

  
ames %>% 
  ggplot(aes(y = SalePrice, x = OverallQual, col = SaleType)) + 
  geom_point() +
  theme_bw()

ames %>% 
  ggplot(aes(y = SalePrice, x = OverallCond, col = SaleType)) + 
  geom_point() +
  theme_bw()


ames %>% 
  ggplot(aes(x = LotArea, y = OverallQual, col = PriceRange)) + 
  geom_point() +
  theme_bw()

ames %>% 
  ggplot(aes(x = GrLivArea, y = OverallQual, col = PriceRange)) + 
  geom_point() +
  theme_bw()

ames %>% 
  ggplot(aes(x = SecFlrSF, y = OverallQual, col = PriceRange)) + 
  geom_point() +
  theme_bw()
  
  
```


```{r}
model.nointeraction = lm(SalePrice ~ ExterCond + sqrt(GarageArea), data=ames_train_)
model.interaction = lm(SalePrice ~ ExterCond:sqrt(GarageArea), data=ames_train_)

anova(model.nointeraction, model.interaction)
```


```{r}
test_interaction = function(feature1, feature2) {
  
  model.nointeraction = lm(SalePrice ~ GrLivArea + feature1 + feature2, data=ames_train)
  model.interaction = lm(SalePrice ~ GrLivArea + feature1:feature2, data=ames_train)
  #print(summary(model.interaction))
  print(anova(model.nointeraction, model.interaction))
}
```
```{r}
#test_interaction(ames_train$OverallCond, ames_train$Neighborhood)
#test_interaction(ames_train$OverallQual, ames_train$Neighborhood)
a = test_interaction(ames_train$Neighborhood, ames_train$OverallCond)

```
```{r}

test_interaction(ames_train$ExterCond, sqrt(ames_train$GarageArea))
```


### Numerical Features
```{r}
library(ggfortify)
```

```{r}
require(gridExtra)
plot_numerical_bands = function(df, x_name, y_name) {
  x = df[,x_name][[1]]
  y = df[,y_name][[1]]
  model = lm(y ~ x, data=df)
  predicted = predict(model)
  
  
  
  g = ggplot(df, aes(x,y)) +
    geom_smooth(method = "lm", se = FALSE, color = "lightgrey") +
    geom_segment(aes(xend = x, yend=predicted), alpha=0.2) +
    geom_point(aes(color=abs(model$residuals))) +
    scale_color_continuous(low = "black", high = "red") +
    guides(color=FALSE) +
    geom_point(aes(y=predicted),shape=1) +
    labs(x = x_name, y = y_name) +
    theme_bw()
  
  # Construction of Confidence Intervals
  
  clower = predict(model, interval='confidence', level=0.95)[,2]
  cupper = predict(model, interval='confidence', level=0.95)[,3]
  
  
  plower = predict(model, interval='prediction', level=0.95)[,2]
  pupper = predict(model, interval='prediction', level=0.95)[,3]
  
  g2 = g + geom_ribbon(aes(ymin = clower, ymax = cupper), alpha=0.3, col = 'red') +
    geom_ribbon(aes(ymin=plower, ymax = pupper), alpha=0.2, col='blue')

  model_name = paste0(y_name,'~',x_name)
  dir.create(paste0("./Linear Model/",model_name))
  
  # Save Scatter Plot
  ggsave(paste0("./Linear Model/",model_name,"/scatter_plot.pdf"), g2)
  
  # Save Diagnostic PLots
  pdf(paste0("./Linear Model/",model_name,"/residual_plots.pdf"))
  par(mfrow=c(2,2))
  plot(model,which=c(1,2,4,6))
  dev.off()
  
  # Save Influence Plot
  pdf(paste0("./Linear Model/",model_name,"/influence.pdf"))
  inf = influencePlot(model)

  dev.off()
  
  
  # plot(x, 
  #      abs(model$residuals), 
  #      xlab = x_name, 
  #      ylab = 'Residual Values Squared', 
  #      main = 'Squared Residuals compared to MSE')
  # abline(h=sqrt(mean(model$residuals^2)), lty = 2, lwd=3, col = 'red')
  # ss = smooth.spline(x, model$residuals^2, cv = TRUE)
  # lines(ss, lwd=2, col='blue')
  # abline(h=mean(ss$y), lwd=2, col='black')
  # legend("topleft", 
  #      c("MSE", "Spline", "Spline Mean"),
  #      lty = c(2, 1, 1), 
  #      col = c("red", "blue", "black"))
}
```



```{r}
require(gridExtra)
plot_cat_bands = function(df, x_name, y_name) {
  x = df[,x_name][[1]]
  y = df[,y_name][[1]]
  model = lm(y ~ x, data=df)
  predicted = predict(model)
  
  
  
  g = ggplot(df, aes(x,y)) +
    geom_smooth(method = "lm", se = FALSE, color = "lightgrey") +
    geom_segment(aes(xend = x, yend=predicted), alpha=0.2) +
    geom_point(aes(color=abs(model$residuals))) +
    scale_color_continuous(low = "black", high = "red") +
    guides(color=FALSE) +
    geom_point(aes(y=predicted),shape=1) +
    labs(x = x_name, y = y_name) +
    theme_bw()
  
  # Construction of Confidence Intervals
  
  clower = predict(model, interval='confidence', level=0.95)[,2]
  cupper = predict(model, interval='confidence', level=0.95)[,3]
  
  
  plower = predict(model, interval='prediction', level=0.95)[,2]
  pupper = predict(model, interval='prediction', level=0.95)[,3]
  
  g2 = g + geom_ribbon(aes(ymin = clower, ymax = cupper), alpha=0.3, col = 'red') +
    geom_ribbon(aes(ymin=plower, ymax = pupper), alpha=0.2, col='blue')

  model_name = paste0(y_name,'~',x_name)
  dir.create(paste0("./Linear Model/",model_name))
  
  # Save Scatter Plot
  ggsave(paste0("./Linear Model/",model_name,"/scatter_plot.pdf"), g2)
  
  # Save Diagnostic PLots
  pdf(paste0("./Linear Model/",model_name,"/residual_plots.pdf"))
  par(mfrow=c(2,2))
  plot(model,which=c(1,2,4,6))
  dev.off()
  
  # Save Influence Plot
  pdf(paste0("./Linear Model/",model_name,"/influence.pdf"))
  inf = influencePlot(model)

  dev.off()
  
  
  # plot(x, 
  #      abs(model$residuals), 
  #      xlab = x_name, 
  #      ylab = 'Residual Values Squared', 
  #      main = 'Squared Residuals compared to MSE')
  # abline(h=sqrt(mean(model$residuals^2)), lty = 2, lwd=3, col = 'red')
  # ss = smooth.spline(x, model$residuals^2, cv = TRUE)
  # lines(ss, lwd=2, col='blue')
  # abline(h=mean(ss$y), lwd=2, col='black')
  # legend("topleft", 
  #      c("MSE", "Spline", "Spline Mean"),
  #      lty = c(2, 1, 1), 
  #      col = c("red", "blue", "black"))
}
```

```{r}

sf_bins = rbin_manual(ames_train, SalePrice, TotalSF, c(1e2,1e3,1e4,1e5,1e6))

plot(sf_bins)
```

```{r}
ames_train %>% 
  mutate(BedroomAbvGr = as.factor(BedroomAbvGr)) %>% 
  group_by(BedroomAbvGr) %>% 
  ggplot(aes(x = GrLivArea, y = SalePrice)) + geom_point() + 
  geom_smooth(method = "lm", se = TRUE, color = "red") +
  facet_wrap(BedroomAbvGr~.)
  
```
```{r}
feature = 'Alley'
ames_train %>% 
  mutate(feature = as.factor(feature)) %>% 
  group_by(feature) %>% 
  ggplot(aes(x = TotalSF, y = SalePrice)) + geom_point() + 
  geom_smooth(method = "lm", se = TRUE, color = "red") +
  #scale_y_continuous(trans='log') +
  facet_wrap(feature)

feature = 'Alley'
ames_train %>% 
  #mutate(feature = as.factor(feature)) %>% 
  group_by(Alley) %>% 
  summarise(SalePrice = mean(SalePrice)) %>% 
  ggplot(aes(x = Alley, y = SalePrice)) + geom_col()
  #geom_smooth(method = "lm", se = TRUE, color = "red") +
  #scale_y_continuous(trans='log') +
  #facet_wrap(feature)
```
```{r}
library(scales)
```



```{r}
plot_pred_by_nom = function(nom_feature, predictor, scale_log = FALSE) {
  
  # x = ames_train[,predictor][[1]]
  # #x2 = ames_train[,nom_feature][[1]]
  # model = lm(ames_train$SalePrice ~ x)
  # predicted = predict(model)
  
  g = ames_train %>% 
    mutate(nom_feature = as.factor(nom_feature)) %>% 
    group_by(nom_feature) %>% 
    ggplot(aes_string(x = predictor, y = 'SalePrice'), environment=environment()) + 
    geom_point(shape=1,alpha=0.4) + 
    geom_smooth(method = "lm", se = TRUE, color = "red", linetype='dashed') +
    # geom_segment(aes(xend = x, yend=predicted), alpha=0.2) +
    # geom_point(aes(color=abs(model$residuals))) +
    # scale_color_continuous(low = "black", high = "red") +
    # guides(color=FALSE) +
    # geom_point(aes(y=predicted),shape=1, alpha=0.2) +
    facet_wrap(nom_feature) +
    labs(title = paste0(predictor,' v Sale Price for given ',nom_feature)) +
    theme_bw()
  
  
  model_name = paste0('SalePrice~',predictor)
  if(scale_log){
    g = g + 
      scale_y_continuous(trans='log') + 
      labs(title=paste0(predictor,' v Log Sale Price for given ',nom_feature))
    
    model_name = paste0('Log_SalePrice~',predictor)
  }
  
  file_name = paste0("./Linear Model/Ordinals/",nom_feature,'/',model_name,'.pdf')
  dir.create(paste0("./Linear Model/Ordinals/",nom_feature))
  # Save Scatter Plot
  ggsave(file_name, g)
  #dev.off()
  print(g)
  
}
```

```{r}
plot_pred_by_nom('MSSubClass',"TotalSF", scale_log=F)
```

```{r}
feature = 'HasPool'
ames_train %>% 
  mutate(feature = as.factor(feature)) %>% 
  #group_by(feature) %>% 
  ggplot(aes(x = TotalSF, y = SalePrice)) + geom_point() + 
  geom_smooth(method = "lm", se = TRUE, color = "red") +
  facet_grid(feature) +
  scale_y_continuous(trans='log')
```

```{r}
feature = 'HasGarage'
ames_train %>% 
  mutate(feature = as.factor(feature)) %>% 
  group_by(feature) %>% 
  ggplot(aes(x = TotalSF, y = SalePrice)) + geom_point() + 
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  facet_wrap(feature) +
  scale_y_continuous(trans='log')
```

```{r}

boxplot(log(ames_train$SalePrice) ~ ames_train$MoSold, outline = FALSE)
boxplot(ames_train$SalePrice ~ ames_train$YrSold, outline = FALSE)
boxplot(log(ames_train$SalePrice) ~ ames_train$YearBuilt, outline = FALSE)
```



```{r}
feature = 'BsmtExposure'
ames_train %>% 
  mutate(feature = as.factor(feature)) %>% 
  group_by(feature) %>% 
  ggplot(aes(x = TotalSF, y = SalePrice)) + geom_point() + 
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  scale_y_continuous(trans='log') +
  facet_wrap(feature)
```
Lets look at Total square footage across different defined sub classes
```{r}
feature = 'MSSubClass'
g = ames_train %>% 
  #filter(feature < 70.0) %>% 
  mutate(feature = as.factor(feature)) %>% 
  group_by(feature) %>% 
  ggplot(aes(x = TotalSF, y = SalePrice)) + geom_point() + 
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  scale_y_continuous(trans='log') +
  facet_wrap(feature)


ggsave("./Linear Model/MsSubClass_logplot.pdf", g)
```



```{r}
feature = 'MSSubClass'
ames_train %>% 
  filter(AfterWW2==1) %>% 
  mutate(feature = as.factor(feature)) %>% 
  group_by(feature) %>% 
  ggplot(aes(x = GrLivArea, y = SalePrice)) + geom_point() + 
  geom_smooth(method = "lm", se = TRUE, color = "red") +
  #geom_spline(color = "blue") +
  #scale_y_continuous(trans='log') +
  facet_wrap(feature)
```


```{r}

ggplot(aes(x, abs(model$residuals)), 
     xlab = x_name, 
     ylab = 'Residual Values Squared', 
     main = 'Squared Residuals compared to MSE')
abline(h=sqrt(mean(model$residuals^2)), lty = 2, lwd=3, col = 'red')
ss = smooth.spline(x, model$residuals^2, cv = TRUE)
lines(ss, lwd=2, col='blue')
abline(h=mean(ss$y), lwd=2, col='black')
```
```{r}
model <- eval(bquote(lm(.(f), data = ames_train)))
```


```{r}
library(broom)

# Steps 1 and 2
d <- lm(mpg ~ hp, data = mtcars) %>% 
       augment()

head(d)

# Steps 3 and 4
ggplot(d, aes(x = hp, y = mpg)) +
  geom_smooth(method = "lm", se = FALSE, color = "lightgrey") +
  geom_segment(aes(xend = hp, yend = .fitted), alpha = .2) +  # Note `.fitted`
  geom_point(aes(alpha = abs(.resid))) +  # Note `.resid`
  guides(alpha = FALSE) +
  geom_point(aes(y = .fitted), shape = 1) +  # Note `.fitted`
  theme_bw()
```



```{r}

# The new line of code


model = lm(log(SalePrice) ~ log(TotalSF), data=ames_train)
predicted = predict(model)

g = ames_train %>% 
  mutate(SalePrice = log(SalePrice), TotalSF = log(TotalSF)) %>% 
  ggplot(aes(x=TotalSF,y=SalePrice)) +
  #geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "lightgrey") +
  geom_segment(aes(xend = TotalSF, yend=predicted), alpha=0.2) +
  geom_point(aes(color=abs(model$residuals))) +
  scale_color_continuous(low = "black", high = "red") +
  guides(color=FALSE) +
  geom_point(aes(y=predicted),shape=1) +
  theme_bw()

# Construction of Confidence Intervals

clower = predict(model, interval='confidence', level=0.95)[,2]
cupper = predict(model, interval='confidence', level=0.95)[,3]


plower = predict(model, interval='prediction', level=0.95)[,2]
pupper = predict(model, interval='prediction', level=0.95)[,3]

g + geom_ribbon(aes(ymin = clower, ymax = cupper), alpha=0.3, col = 'red') +
  geom_ribbon(aes(ymin=plower, ymax = pupper), alpha=0.2, col='blue')
```
Definitely a logarithimic relationship...


```{r}
qqnorm(model$residuals)
qqline(model$residuals)
```


```{r}
ames_train %>% 
  ggplot(aes(x = TotalSF, y = SalePrice)) +
  #geom_point() +
  geom_hex(bins=55)

```


### Automating Variable Selection Process

According to Akaike Criterion
```{r}
library(MASS)
```

```{r}
model.empty = lm(SalePrice ~ 1, data = train) #The model with an intercept ONLY.
model.full = lm(SalePrice ~ ., data = train) #The model with ALL variables.
scope = list(lower = formula(model.empty), upper = formula(model.full))
```

```{r}
forwardAIC = step(model.empty, scope, direction = "forward", k = 2)
backwardAIC = step(model.full, scope, direction = "backward", k = 2)
bothAIC.empty = step(model.empty, scope, direction = "both", k = 2)
bothAIC.full = step(model.full, scope, direction = "both", k = 2)
```

```{r}
summary(bothAIC.empty)
```

```{r}
plot(bothAIC.full)
```


```{r}
influencePlot(bothAIC.full)
```
### Prediction Function
```{r}
make_pred = function(model, data) {
  data = data[,attr(model$terms, 'term.labels')]
  pred.band = predict(model, data, interval='prediction')
}
```



```{r}
sapply(num_train[rownames(inf),attr(bothAIC.full$terms, 'term.labels')],exp) %>% data.frame()
```

```{r}
summary(bothAIC.empty)
```



### Discrete Numerical Features EDA

```{r}
features = c('YearBuilt','YearRemodAdd','BsmtFullBath','BsmtHalfBath','Full','SalePrice')

train_dis = train[,features]
#test_dis = test[,features]

```

```{r}
clean_data_str = function(date_str) {
  mdy = mdy(date_str)
  ymd = ymd(date_str)
  
  if(is.na(mdy)){
    if(is.na(ymd)){
      return(NA)
    }
    else {return(ymd)}
  }
  else {return(mdy)}
}

```

```{r}
train_dis %>% ggplot(aes(x = YearBuilt, y = SalePrice)) + geom_point()
```

Split into test and train
```{r}
library(tidyverse)

x = model.matrix(SalePrice ~ ., ames)[, -1]
y = ames$SalePrice


set.seed(0)
train = sample(1:nrow(x), 7*nrow(x)/10)
y.test = y[-train]


```



```{r}
library(caret)
set.seed(0)
grid = 10^seq(5, -2, length = 100)
```

```{r}

cv.lasso.out = cv.glmnet(x[train,],y[train], lambda=grid, alpha=1, nfolds=10)
plot(cv.lasso.out)

```

```{r}
bestlambda.lasso = cv.lasso.out$lambda.min
bestlambda.lasso
log(bestlambda.lasso)
```

```{r}
lasso.bestlambdatrain = predict(lasso.models.train, s = bestlambda.lasso, newx = x[-train,])

mean((lasso.bestlambdatrain - y.test)^2)
```
```{r}
tmp_coeffs <- coef(cv.glmnet.fit, s = "lambda.min")
data.frame(name = tmp_coeffs@Dimnames[[1]][tmp_coeffs@i + 1], coefficient = tmp_coeffs@x)
```

```{r}
coefs.lasso = coef(cv.lasso.out)
```


```{r}
str(coefs.lasso)
```

```{r}
coefs.lasso@x
```

```{r}
gam1 = mgcv::gam(SalePrice ~ s(TotalSF, bs='ps', sp=0.6) + s(KitchenAbvGr, bs='ps', sp=0.6), data=ames_train)
```

```{r}
### GAM example using mgcv

library(mgcv)
library(ggplot2)
# fake data
n <- 50
sig <- 2
dat <- gamSim(1,n=n,scale=sig)

# P-spline smoothers (with lambda=0.6) used for x1 and x2; x3 is parametric.
b1 <- mgcv::gam(y ~ s(x1, bs='ps', sp=0.6) + s(x2, bs='ps', sp=0.6) + x3, data = dat)
summary(b1)
plot(b1)


# plot the smooth predictor function for x1 with ggplot to get a nicer looking graph
p <- predict(b1, type="lpmatrix")
beta <- coef(b1)[grepl("x1", names(coef(b1)))]
s <- p[,grepl("x1", colnames(p))] %*% beta
ggplot(data=cbind.data.frame(s, dat$x1), aes(x=dat$x1, y=s)) + geom_line()


# predict
newdf <- gamSim(1,n=n,scale=sig)
f <- predict(b1, newdata=newdf)


# select smoothing parameters with REML, using P-splines
b2 <- mgcv::gam(y ~ s(x1, bs='ps') + s(x2, bs='ps') + x3, data = dat, method="REML")

# select variables and smoothing parameters
b3 <- mgcv::gam(y ~ s(x0) + s(x1) + s(x2) + s(x3) , data = dat, method="REML", select=TRUE)

# loess smoothers with the gam package (restart R before loading gam)
library(gam)
b4 <- gam::gam(y ~ lo(x1, span=0.6) + lo(x2, span=0.6) + x3, data = dat)
summary(b4)
```

```{r}
feature = 'OverallQual'

df = ames_train %>% group_by(AfterWW2,HouseStyle) %>% summarise(price_med = median(SalePrice), n =n())

ames_train %>% 
  ggplot(aes(x = factor(HouseStyle), y = SalePrice)) + geom_boxplot() +
  geom_text(data=df, color='red',size=4,aes(y=price_med*0.8, label=n)) +
  facet_grid(.~factor(AfterWW2)) +
  labs() +
  #scale_y_continuous(trans='log') +
  theme_bw()

```

```{r}
model.fire = lm(log(SalePrice) ~ sqrt(GrLivArea) + sqrt(TotalBsmtSF) + OverallCond + AfterWW2 + Fireplaces + IsNew + OverallQual, data = ames)
summary(model.fire)
plot(model.fire)
inf = influencePlot(model.fire)
ames[rownames(inf),]
```

```{r}
model = lm(log(SalePrice) ~ TotalBsmtSF + GrLivArea, data = ames)
 # Create a vector of gradually-changing colors, with one entry for each data # point
the.colors <- rainbow(n = nrow(df))
# For each data point, see how it ranks according to X2, from smallest (1)
# to largest
the.ranks <- rank(ames$GrLivArea)
# Plot residuals vs. X1, colored according to X2 Defining the color and rank
# vectors makes this next line a bit less mysterious, but it's not
# necessary; this could all be a one-liner.
plot(ames$TotalBsmtSF, residuals(model), pch = 19, col = the.colors[the.ranks], ylab = "Residuals")
```

```{r}
hist(sqrt(ames$TotalBsmtSF))
hist(ames$TotalBsmtSF)
hist(log(ames$SalePrice))
```



```{r}
library(tidyverse)
model.test = lm(SalePrice ~ TotalSF + AfterWW2, data = ames)
res = data.frame(cbind(ames$AfterWW2, residuals(model.test)))
colnames(res) = c('AfterWW2','Residuals')
ggplot(data = res, aes(x = factor(AfterWW2), y = Residuals)) + geom_boxplot()
ggplot(data = res, aes(x=Residuals)) + geom_density(alpha=.2, fill="#FF6666") + facet_grid(res$AfterWW2, scales = 'free_y')
```

```{r}
ames %>% 
  filter(YearBuilt > 1990) %>% 
  group_by(YearBuilt,CentralAir) %>% 
  tally() %>% 
  ggplot(aes(x = YearBuilt, y=n, fill=CentralAir)) + geom_col()
```
```{r}
ames %>% 
  filter(YearBuilt > 1990) %>% 
  group_by(YearBuilt,MSSubClass) %>% 
  tally() %>% 
  ggplot(aes(x = YearBuilt, y=n, fill=factor(MSSubClass))) + geom_col()
```

```{r}
hist(ames$SalePrice)
```
```{r}

```


```{r}
b = c(-Inf, 1.2e5, 2e5,5e5,Inf)
names = c('Cheap','Average','Expensive','Very Expensive')
bins = cut(ames$SalePrice, breaks = b, labels = names)

ames %>% 
  cbind(., bins) %>% 
  filter(YearBuilt > 1930) %>% 
  group_by(YearBuilt,bins) %>% 
  tally() %>% 
  ggplot(aes(x = YearBuilt, y=n, fill=factor(bins))) + geom_col()

```

```{r}
ames %>% 
  group_by(YrSold,YearBuilt) %>% 
  tally() %>% 
  ggplot(aes(x = YrSold, y=n, fill=YearBuilt)) + geom_col(position='fill')
```

```{r}
model.test = lm(log(SalePrice) ~ sqrt(TotalSF)*MSSubClass,data=ames )
summary(model.test)
```

```{r}
library(AppliedPredictiveModeling)
transparentTheme(trans = .4)
library(caret)
plotSubset <- data.frame(scale(ames[, c("SalePrice", "MSSubClass")])) 
xyplot(SalePrice ~ MSSubClass,
       data = ames,
       groups = ames$AfterWW2, 
       auto.key = list(columns = 2)) 
```

```{r}
transformed <- spatialSign(plotSubset)
transformed <- as.data.frame(transformed)
xyplot(SalePrice ~ MSSubClass, 
       data = transformed, 
       groups = ames$AfterWW2, 
       auto.key = list(columns = 2)) 
```

```{r}
ames_train %>% 
  group_by(MSSubClass,AfterWW2) %>% 
  ggplot(aes(x = factor(MSSubClass), y=SalePrice)) + geom_boxplot() +facet_grid(('AfterWW2'))

ames_train %>% 
  group_by(MSSubClass,HouseStyle) %>% 
  ggplot(aes(x = factor(MSSubClass), y=SalePrice)) + geom_boxplot() +facet_wrap(('HouseStyle'))
```

```{r}
test_interaction(ames_train$ExterCond, ames_train$GarageArea)
```

