---
title: "Missingness in Ames"
output:
  pdf_document: default
  html_document:
    df_print: paged
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message=FALSE)
```


```{r}
library(Hmisc)
library(psych)
library(kknn)
library(mice)
library(VIM)
library(tidyverse)
library(lattice)
```

```{r}
ames = read_csv('./data/train.csv')
ames = ames[,-1]
```

## Summary
1. Summary Metrics for Numeric Features
```{r}
# Take out Id column
ames %>% select_if(is.double) %>% describe()
```

2. How many observations with at least one missing value?
```{r}
sum(!complete.cases(ames))
mean(!complete.cases(ames)) * 100
```
All of them. 


3. How many cells have missing values?
```{r}
sum(is.na(ames))
mean(is.na(ames)) * 100
```
Almost 6% of the dataset is missing

4. Different combinations of missingness in the dataset
```{r}
ames_aggr = ames %>% aggr(., col=mdc(1:2), numbers=TRUE, combine=FALSE, sortVars=TRUE, labels=names(ames), cex.axis=0.8, gap=2, ylab=c("Proportion of missingness\n among numeric features","Missingness Pattern"))
```
Seems a lot of missingness has to do with homes with missing pool and other sundry details. 
Lets filter those out to get a better sense of overall missingness. 

```{r}
ames_aggr = ames %>% 
  dplyr::select(.,-c('GarageCond','Fence','Alley','MiscFeature',
                     'ScreenPorch','FireplaceQu','PoolArea','PoolQC')) %>% 
  aggr(., col=mdc(1:2), numbers=TRUE, combine=FALSE, sortVars=TRUE, labels=names(ames), cex.axis=0.8, gap=2, ylab=c("Proportion of missingness\n among numeric features","Missingness Pattern"))
```
## Imputation

1. Lot Frontage: Linear Feet of street connected to property
```{r}
densityPlot(ames$LotFrontage)
```

2. Random Imputation
```{r}
ames.lotF = Hmisc::impute(ames$LotFrontage, "random")
plot(density(ames.lotF), col = "red",
     main = "Random Imputation for Age")
lines(density(ames$LotFrontage, na.rm = TRUE), col = "blue")
legend("topright", c("Original", "Random Imputed"),
       col = c("blue", "red"), lwd = 1)
```
Though it looks pretty good, we can't be sure we're assigning correct lot frontages to the correct houses (if even applicable). A more defensible procedure would be to use k-means imputation or multiple imputation for doing statistical prediction. 

3. Multiple Imputation
```{r}
# Examine the correlation of LotFrontage with other variables
numeric_feat = ames %>% dplyr::select_if(is.double)
cor_numeric_pairwise = cor(numeric_feat, use='pairwise.complete.obs')
cor_numeric_complete = cor(numeric_feat, use='complete.obs')

```
```{r}
cor_numeric_pairwise[,'LotFrontage'] %>% sort()
```
High correlation with 1st floor square footage (1stFlrSF), Lot area (LotArea), ground living area (GrLivArea)
Let's throw in sale price for an evalution of how price would affect the reporting of this feature
Does having a garage (GarageArea) or number car garage (GarageCars) affect reporting levels?
Does overall quality (OverallQual) of the home affect reporting?
Lets see some margin plots to evaluate the type of missingness in this feature
```{r}

marginplot(ames[, c("1stFlrSF","LotFrontage")], col = mdc(1:2), cex.numbers = 1.2, pch = 19)
# No change
marginplot(ames[, c("LotArea","LotFrontage")], col = mdc(1:2), cex.numbers = 1.2, pch = 19)
# No change
marginplot(ames[, c("GrLivArea","LotFrontage")], col = mdc(1:2), cex.numbers = 1.2, pch = 19)
# Almost no change
marginplot(ames[, c("SalePrice","LotFrontage")], col = mdc(1:2), cex.numbers = 1.2, pch = 19)
# Median sale price of homes without reported lot frontage is higher than those with reported
marginplot(ames[, c("GarageArea","LotFrontage")], col = mdc(1:2), cex.numbers = 1.2, pch = 19)
# Size of garages with no LotFrontage reported tend to cluster between 400 and 600 sf
marginplot(ames[, c("GarageCars","LotFrontage")], col = mdc(1:2), cex.numbers = 1.2, pch = 19)
# Number car garage has no effect
marginplot(ames[, c("OverallQual","LotFrontage")], col = mdc(1:2), cex.numbers = 1.2, pch = 19)
# OverallQual has no effect
```


```{r}
ames = ames %>% rename('SecFlrSF'='2ndFlrSF','ThreePorch'='3SsnPorch')
```

Correlation with potential predictors
```{r}
features = c('FirstFlrSF','LotArea','GrLivArea','SalePrice','GarageArea')
cor(ames[,features])
```


Lets evaluate the relationship as MCAR and use the features for which LotFrontage exhibits MAR behaviour as predictors for the missing values
```{r}
# Function mice() in mice package is a Markov Chain Monte Carlo (MCMC) method that uses 
# correlation structure of the data and imputes missing values for each incomplete 
# variable m times by regression of incomplete variables on the other variables iteratively. 
imp = mice(ames, m=20, printFlag=FALSE, maxit = 40, seed=2525)
```
```{r}
# The output imp contains m=5 completed datasets. Each dataset can be analysed
# using function with(), and including an expression for the statistical analysis approach
# we want to apply for each imputed dataset as follows
fit.mi = with(data=imp, exp = lm(LotFrontage ~ sqrt(FirstFlrSF) + sqrt(LotArea)))
combFit = mice::pool(fit.mi) 
summary(combFit)
mice::pool.r.squared(fit.mi)
```



Check for implausibility in imputations (LF < 0)
```{r}
sum(imp$imp$LotFrontage <10)
```

# Under MAR assumption, the densities should exhibit simlarily (though not be identical)
```{r}
densityplot(imp, SalePrice ~ LotFrontage)

```

testing the fit
```{r}
fit1 = with(data = imp, expr = lm(LotFrontage ~ SalePrice + FirstFlrSF + LotArea))
fit2 = with(data = imp, expr = lm(LotFrontage ~ SalePrice + FirstFlrSF + GrLivArea + LotArea))
# Wald test
stat = mice::pool.compare(fit2, fit1)
# P-value of the test
stat$p
```
```{r}
summary(combFit)
```

```{r}
mice::complete(imp, "long")
```

